---
title: "French given names per year per department"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "October, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the [given names data
set of INSEE](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2017_txt.zip), answer some of the following questions:

- First name frequency evolves along time?
- What can we say about "the french version of your name here" (for each state, all the country)?
- Is there some sort of geographical correlation with the data?
- Which state has a larger variety of names along time?
- _your own question_ (be creative)

You need to use the _dplyr_ and _magrittr_ for this analysis. Read in R with this code. Note that you might need to install the `readr` package with the appropriate command.

```{r}
file = "dpt2017_txt.zip"
if(!file.exists(file)){
  download.file("https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2017_txt.zip",
	destfile=file)
}
unzip(file)
```

```{r}
library(readr);
df <- read_tsv("dpt2017.txt", locale = locale(encoding = "ISO-8859-1"), col_types=cols(
  sexe = col_integer(),
  preusuel = col_character(),
  annais = col_character(),
  dpt = col_character(),
  nombre = col_integer()
));
df;
```


```{r}
library(dplyr)
library(ggplot2)
```

Loaded library.

Trying to answer the following question:
- Does First name frequency evolve along time?

```{r}
df %>% group_by(preusuel,annais) %>% summarize(count = sum(nombre)) %>% filter(preusuel != "_PRENOMS_RARES") %>% arrange(as.integer(annais)) %>% ggplot(aes(x=as.integer(annais), y=count)) + geom_point() + geom_line()
```

Clearly, the frequency of all names is more homogeneous now.

Second question: what can you say about Francis/Francois?

```{r}
df %>%filter(preusuel == "FRANCIS") %>% group_by(preusuel,annais) %>% summarize(count = sum(nombre)) %>%  arrange(annais) %>% ggplot(aes(x=as.integer(annais), y=count)) + geom_point() + geom_line()
```

Francis is a somewhat rare name in France, with a peak of popularity in the 50's (perhaps due to Francis Ford Coppola's "The Grandfather" movie?)

```{r}
df %>% filter(preusuel == "FRANCIS") %>% group_by(dpt,annais) %>% summarize(count = sum(nombre)) %>%
arrange(annais) %>% ggplot(aes(x=as.integer(dpt), y=count)) + geom_point()
```

The name "Francis" fairly more frequent in 2 state departments

Which state has a larger variety of names along time?

```{r}
df %>% group_by(dpt,annais) %>% summarize(count = n()) %>% arrange(annais) %>% ggplot(aes(x=as.integer(annais), y=count, color(dpt))) + geom_point()
```

```{r}
df %>% group_by(dpt,annais) %>% summarize(count = n()) %>% arrange(annais) %>% ggplot(aes(x=as.integer(annais), y=count, color(dpt))) + geom_point()
```

